---
title: 'Case study: How does a bike-share navigate speedy success?'
author: "Zelimkhan"
date: '2025-05-31'
output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

::: {style="text-align: justify"}
This is my capstone project of the Google Data Analytics Professional
Certificate. In this case study, I work for a fictional company,
Cyclistic, along with some key team members. In order to answer the
business questions, follow the steps of the data analysis process: Ask,
Prepare, Process, Analyze, Share, and Act.
:::

# Scenario

::: {style="text-align: justify"}
I am a junior data analyst working on the marketing analyst team at
Cyclistic, a bike-share company in Chicago. The director of marketing
believes the company’s future success depends on maximizing the number
of annual memberships. Therefore, my team wants to understand how casual
riders and annual members use Cyclistic bikes differently. From these
insights, my team will design a new marketing strategy to convert casual
riders into annual members. But first, Cyclistic executives must approve
my recommendations, so they must be backed up with compeling data
insights and professional data visualizations.
:::

### Characters and teams

::: {style="text-align: justify"}
-   **Cyclistic:** A bike-share program that features more than 5,800
    bicycles and 600 docking stations. Cyclistic sets itself apart by
    also offering reclining bikes, hand tricycles, and cargo bikes,
    making bike-share more inclusive to people with disabilities and
    riders who can’t use a standard two-wheeled bike. The majority of
    riders opt for traditional bikes; about 8% of riders use the
    assistive options. Cyclistic users are more likely to ride for
    leisure, but about 30% use the bikes to commute to work each day.
-   **Lily Moreno:** The director of marketing and your manager. Moreno
    is responsible for the development of campaigns and initiatives to
    promote the bike-share program. These may include email, social
    media, and other channels.
-   **Cyclistic marketing analytics team:** A team of data analysts who
    are responsible for colecting, analyzing, and reporting data that
    helps guide Cyclistic marketing strategy. You joined this team six
    months ago and have been busy learning about Cyclistic’s mission and
    business goals—as wel as how you, as a junior data analyst, can help
    Cyclistic achieve them.
-   **Cyclistic executive team:** The notoriously detail-oriented
    executive team wil decide whether to approve the recommended
    marketing program.
:::

### About the company

:::: {style="text-align: justify"}
In 2016, Cyclistic launched a successful bike-share oering. Since then,
the program has grown to a eet of 5,824 bicycles that are geotracked and
locked into a network of 692 stations across Chicago. The bikes can be
unlocked from one station and returned to any other station in the
system anytime. <br>

Until now, Cyclistic’s marketing strategy relied on building general
awareness and appealing to broad consumer segments. One approach that
helped make these things possible was the flexibility of its pricing
plans: single-ride passes, full-day passes, and annual memberships.
Customers who purchase single-ride or full-day passes are referred to as
casual riders. Customers who purchase annual memberships are Cyclistic
members.<br>

Cyclistic’s finance analysts have concluded that annual members are much
more protable than casual riders. Although the pricing flexibility helps
Cyclistic attract more customers, Moreno believes that maximizing the
number of annual members will be key to future growth. Rather than
creating a marketing campaign that targets all-new customers, Moreno
believes there is a solid opportunity to convert casual riders into
members. She notes that casual riders are already aware of the Cyclistic
program and have chosen Cyclistic for their mobility needs.<br>

Moreno has set a clear goal: Design marketing strategies aimed at
converting casual riders into annual members. In order to do that,
however, the team needs to beer understand how annual members and casual
riders differ, why casual riders would buy a membership, and how digital
media could affect their marketing tactics. Moreno and her team are
interested in analyzing the Cyclistic historical bike trip data to
identify trends.

Three questions will guide the future marketing program: <br> 1. How do
annual members and casual riders use Cyclistic bikes dierently? <br> 2.
Why would casual riders buy Cyclistic annual memberships? <br> 3. How
can Cyclistic use digital media to inuence casual riders to become
members? <br>

Moreno has assigned you the first question to answer: How do annual
members and casual riders use Cyclistic bikes dierently?

# Step 1. Ask

<br>

| № | Type | Questions | Answers |
|---------------|---------------|---------------|---------------------------|
| <br> | <br> | <br> | <br> |
| 1 | Guiding questions | What is the problem you are trying to solve? | The problem is understanding how annual members and casual riders use Cyclistic bikes differently. This insight will help the marketing team design strategies to convert casual riders into annual members, which are more profitable for the company. |
| <br> | <br> | <br> | <br> |
| 2 | Guiding questions | How can your insights drive business decisions? | By identifying key differences in usage patterns (e.g., ride duration, frequency, preferred days/times), the marketing team can tailor campaigns to highlight the benefits of annual memberships for casual riders. This could include targeted promotions, personalized messaging, or incentives based on their behavior. |
| <br> | <br> | <br> | <br> |
| 3 | Key tasks | Identify the business task | Analyze historical bike trip data to compare usage patterns between casual riders and annual members, and provide actionable recommendations to convert casual riders into members. |
| <br> | <br> | <br> | <br> |
| 4 | Key tasks | Consider key stakeholders | 1) Lily Moreno (Director of Marketing): Needs data-driven insights to justify marketing strategies. <br> 2) Cyclistic Executive Team: Requires compelling evidence to approve recommendations. <br> 3) Marketing Analytics Team: Relies on accurate analysis to guide campaign design. |
| <br> | <br> | <br> | <br> |
| 5 | Deliverable | A clear statement of the business task | Analyze Cyclistic's historical bike-share data to identify differences in usage patterns between casual riders and annual members. The goal is to provide actionable insights that will inform a marketing strategy aimed at converting casual riders into annual members, thereby increasing profitability. |

<br>

# Step 2. Prepare

::: {style="text-align: justify"}
### Guiding Questions & Responses <br>

1.  Where is your data located? <br>

The data is publicly available in this two datasets: <br> \* [Divvy 2019
Q1](https://docs.google.com/spreadsheets/d/1uCTsHlZLm4L7-ueaSLwDg0ut3BP_V4mKDo2IMpaXrk4/template/preview?resourcekey=0-dQAUjAu2UUCsLEQQt20PDA#gid=1797029090)<br>
\* [Divvy 2020
Q1](https://docs.google.com/spreadsheets/d/179QVLO_yu5BJEKFVZShsKag74ZaUYIF6FevLYzs3hRc/template/preview#gid=640449855)<br>

2.  How is the data organized? <br>

The datasets are organized as quarterly google sheets files, with each
row representing a bike trip and columns including: <br> 1 ) ride_id,
rideable_type (bike type) <br> 2 ) started_at, ended_at (timestamps)
<br> 3 ) member_casual (user type) <br> 4 ) start_station_id,
end_station_id (location data) <br>

3.  Are there issues with bias or credibility in this data? Does your
    data ROCCC? <br>

1 ) Reliable: Yes, sourced from a licensed provider (Motivate
International Inc.). <br> 2 ) Original: First-party system-generated
(Cyclistic’s operational system) operational data. <br> 3 )
Comprehensive: Covers all trips during the specified quarters, including
key fields. <br> 4 ) Current: Q1 2019 and Q1 2020 are accurate within
their timeframes. <br> 5 ) Cited: Licensed and publicly cited via
[Divvy's data agreement](https://divvybikes.com/data-license-agreement)
<br> 6 ) Potential Bias: Excludes personally identifiable information
(PII), so demographic analysis is limited. <br>

4.  How are you addressing licensing, privacy, security, and
    accessibility? <br>

1 ) Licensing: Complies with Divvy's public data use license. <br> 2 )
Privacy: No percenal information concluded. <br> 3 ) Security: Data
stored securely in google sheets. <br> 4 ) Accessibility: Data in google
sheets format ensures compatibility with Exel, R or SQL. <br>

5.  How did you verify the data's integrity? <br>

Data integrity was checked by: <br> 1 ) Removing duplicates and
incomplete records. <br> 2 ) Ensuring start times precede end times.
<br> 3 ) Filtering out trips with negative or zero durations. <br> 4 )
Verifying consistent labeling of user types. <br> 5 ) Running summary
statistics to confirm data completeness and logical consistency. <br>

6.  How does it help you answer you questions? <br>

This data allows analysis of how casual riders and annual members use
Cyclistic bikes differently by providing trip-level details such as ride
length, frequency, timing, and bike type preferences. These insights are
crucial for designing targeted marketing strategies.

7.  Are there any problems with the data? <br>

Yes: <br> 1 ) Column names are inconsistent between the two datasets
(e.g., usertype in 2019 vs member_casual in 2020). <br> 2 ) The data
does not include demographic or payment information, limiting deeper
customer profiling. Seasonal and external factors are not directly
captured but may affect usage. The dataset size requires careful
handling to avoid memory issues. <br>

### Key Tasks

| № | Task | Details |
|------------------|---------------------|---------------------------------|
| <br> | <br> | <br> |
| 1 | Download data | Downloaded Q1 2019 and Q1 2020 google sheets files. |
| <br> | <br> | <br> |
| 2 | Store appropriately | Files stored in google drive folder [Source data](https://drive.google.com/drive/folders/1BO4tpPfs25t5z5Sq2a-neWuLGSlcVdyR). |
| <br> | <br> | <br> |
| 3 | Identify how it’s organized | The datasets are organized as quarterly google sheets files, with each row representing a bike trip and columns including: <br> 1 ) ride_id, rideable_type (bike type) <br> 2 ) started_at, ended_at (timestamps) <br> 3 ) member_casual (user type) <br> 4 ) start_station_id, end_station_id (location data) <br> |
| <br> | <br> | <br> |
| 4 | Sort and filter the data | Previewed and confirmed need to normalize column names and filter out invalid rows (e.g., negative ride_length). |
| <br> | <br> | <br> |
| 5 | Determine credibility of data | ... |
| <br> | <br> | <br> |

### Deliverable: Description of All Data Sources Used <br>

Data Sources Used: <br> This project utilizes two publicly available
datasets from the Chicago Divvy bike-share system:<br>
Divvy_Trips_2019_Q1:
[Link](https://docs.google.com/spreadsheets/d/1U-zZ-Kpvs4RNjsL4q2W32h7BdjzENRiAPSkeKPVSEN0/edit?gid=1797029090#gid=1797029090).
<br> Covers January–March 2019.<br> Key fields: trip_id, start_time,
end_time, bikeid, from_station_name, to_station_name, usertype.<br>

Divvy_Tripdata_2020_Q1:
[Link](https://docs.google.com/spreadsheets/d/18vpR9boxiVwoI2k0Pu2Sm6ghfJ_KJAtOozjTCjRLjac/edit?gid=640449855#gid=640449855).
<br> Covers January–March 2020. <br> Key fields: ride_id, started_at,
ended_at, rideable_type, member_casual, start_station_name,
end_station_name. <br>

These datasets were sourced from the official Divvy system and released
by Motivate International Inc.. The data has been vetted for public use
and adheres to privacy requirements (no personally identifiable
information).<br> Licensing information: Divvy License Agreement
:::

# Step 3. Prepare

```{r}
# Instaling neccesary packages
install.packages("googlesheets4")
install.packages("googledrive")
install.packages("tidyverse")
library(googlesheets4)
library(googledrive)
library(tidyverse)
```

```{r}
# Reading data
trip_2019 <- read_sheet("https://docs.google.com/spreadsheets/d/1U-zZ-Kpvs4RNjsL4q2W32h7BdjzENRiAPSkeKPVSEN0/edit?gid=1797029090#gid=1797029090")
trip_2020 <- read_sheet("https://docs.google.com/spreadsheets/d/18vpR9boxiVwoI2k0Pu2Sm6ghfJ_KJAtOozjTCjRLjac/edit?gid=640449855#gid=640449855")
```

```{r}
trip_2019

colnames(trip_2019)
colnames(trip_2020)
```

```{r}
trip_2020
```

::::


