---
title: "1.2. apply1"
author: "Zelimkhan Gudiev"
date: "2022-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install the necessary packages and libraries
```{r}
library(ggplot2)
library(dplyr)
library(readxl)

```

## Upload data
```{r}
getwd()
setwd("C:/Users/GudievZK/Desktop/GitHub/DF/")
# setwd("/Users/zelimkhan/Desktop/Data/GitHub/DF/")#

data("diamonds")

```
## Find the minimum value in three dimensions

```{r}
# min_size <- c()
# for(i in 1:nrow(diamonds)) {
#   min_size <- c(min_size, min(diamonds[i, 8:10]))
# }

min_size <- numeric(nrow(diamonds))
for (i in 1:nrow(diamonds)) {
  min_size[i] <- min(diamonds[i, 8:10])
}
```

Циклы в R работают очень медленно и поэтому лучше использовать альтернативные варианты с применением функций семейства apply

## apply
```{r}
apply(diamonds[, 8:10], 1, min)

```
## Apply Functions Over Array Margins\
**Description**\
Returns a vector or array or list of values obtained by applying a function to margins of an array or matrix.\
**Usage**\
apply(X, MARGIN, FUN, ..., simplify = TRUE)\
**Arguments:**\
**X** -	an array, including a matrix.\
**MARGIN** - a vector giving the subscripts which the function will be applied over. E.g., for a matrix 1 indicates rows, 2 indicates columns, c(1, 2) indicates rows and columns. Where X has named dimnames, it can be a character vector selecting dimension names.\
**FUN**	- the function to be applied: see ‘Details’. In the case of functions like +, %*%, etc., the function name must be backquoted or quoted.\
**...** - optional arguments to FUN.\
**simplify** - a logical indicating whether results should be simplified if possible.\






















