---
title: "1.2. apply1"
author: "Zelimkhan Gudiev"
date: "2022-11-14"
output: html_document
---

В этом уроке мы:
узнаем, что использование циклов for в R может значительно усложнить нашу жизнь;
познакомимся с функцией apply;
решим несколько практических задач, используя apply.

Скрипт занятия доступен по ссылке <https://stepic.org/media/attachments/course/724/apply1.r>.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install the necessary packages and libraries
```{r}
library(ggplot2)
library(dplyr)
library(readxl)
```

## Upload data
```{r}
getwd()
# setwd("C:/Users/GudievZK/Desktop/GitHub/DF/")
setwd("/Users/zelimkhan/Desktop/Data/GitHub/DF/")#

data("diamonds")

```
## Find the minimum value in three dimensions

```{r}
# min_size <- c()
# for(i in 1:nrow(diamonds)) {
#   min_size <- c(min_size, min(diamonds[i, 8:10]))
# }

min_size <- numeric(nrow(diamonds))
for (i in 1:nrow(diamonds)) {
  min_size[i] <- min(diamonds[i, 8:10])
}
```

Циклы в R работают очень медленно и поэтому лучше использовать альтернативные варианты с применением функций семейства apply

## apply
```{r}
min_size2 <- apply(diamonds[, 8:10], 1, min)

```
## Apply Functions Over Array Margins\
**Description**\
Returns a vector or array or list of values obtained by applying a function to margins of an array or matrix.\
**Usage**\
apply(X, MARGIN, FUN, ..., simplify = TRUE)\
**Arguments:**\
**X** -	an array, including a matrix.\
**MARGIN** - a vector giving the subscripts which the function will be applied over. E.g., for a matrix 1 indicates rows, 2 indicates columns, c(1, 2) indicates rows and columns. Where X has named dimnames, it can be a character vector selecting dimension names.\
**FUN**	- the function to be applied: see ‘Details’. In the case of functions like +, %*%, etc., the function name must be backquoted or quoted.\
**...** - optional arguments to FUN.\
**simplify** - a logical indicating whether results should be simplified if possible.\

```{r}
d <- matrix(rnorm(30), nrow = 5)

s <- apply(d, MARGIN = 1, FUN = sd)
```
Function apply returned the vector. The length this vector is equal to the length of the matrix's "d" row.

If we change the MARGIN argument from 1 to 2, we will get a vector whose length is equal to the length of the matrix's "d" column

```{r}
s <- apply(d, MARGIN = 2, FUN = sd)
```
s is vector
```{r}
is.vector(s)
typeof(s)
```


## Step 5 of 9
В переменной my_df сохранен dataframe с произвольным числом количественных переменных. При помощи функции apply найдите максимальное значение в каждой строке. Сохраните результат (вектор максимальных значений) в переменную row_max.
```{r}
my_df <- matrix(rnorm(100), ncol = 10)
```


```{r}
row_max <- apply(my_df, 1, max)
```

## Step 6 of 9
В переменной my_df сохранен dataframe с произвольным числом количественных переменных. Рассчитайте медиану для всех столбцов с количественными переменными. В переменную col_median сохраните вектор полученных значений. 

```{r}
col_median <- apply(my_df, MARGIN = 2, FUN = median)
```

## Step 7 of 9
```{r}
my_range <- apply(d, MARGIN = 2, FUN = range)
my_range
```

In this case function apply returned the matrix. In case when applying a function returns more than one value and these values are equal for every margins, the apply function will return a matrix.\

## Step 8 of 9

```{r}
outliers_count <- function(x) {
  outliers <- x[abs(x - mean(x)) > 2 * sd(x)]
  if (length(outliers) > 0) {
    return(outliers)
  } else {
    return("There are no outliers")
  }
}
```


```{r}
iris_num <- iris[, 1:4]
```

```{r}
apply(iris_num, 2, outliers_count)
```

In this case function apply returned the list. In case when applying a function returns values different length or type for every margins, the apply function will return a list.\

## Step 9 of 9






