---
title: "DA 54. Analysis of changes to the SPGZ in 2025 in terms of grounds"
author: "Zelimkhan Gudiev"
date: "2025-12-17"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
  word_document:
    toc: true
    toc_depth: '5'
  pdf_document:
    toc: true
    toc_depth: '5'
---

::: {style="text-align: justify"}

# Install packages and attach libraries
```{r setup, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
# install.packages("psych")
# install.packages("lubridate")
# install.packages("purrr")
# install.packages("GGally")
# install.packages("skimr")
# install.packages("Writexl")
# install.packages("corrplot")
library(psych)
library(googlesheets4)
library(tidyr)
library(dplyr)
library(janitor)
library(ggplot2)
library(lubridate)
library(purrr)
library(GGally)
library(readxl)
library(skimr)
library(writexl)
library(corrplot)
```

# **1. Business task, context and scope of work**

## **1.1. Key tasks and deliverable(s)**

### **1.1.1. Business task**

Determine the number of changes made to the SPGZ in 2025 for each of possible scenarios (cases) for the subsequent determination of frequency of scenarios (cases) in which customers receive notifications (in accordance with "[EAIST_JB-8765](https://pm.mos.ru/issue/EAIST_JB-8765) Реализовать уведомления пользователям при изменении позиций КПГЗ/СПГЗ, которые были запланированы в лотах") and scenarios (cases) that entail the need for customers to make changes to the lot specifications. 

**Notes**

* The assignment to carry out of this task was issued at the "[Meet-145](https://youtrack.gbuac.moscow/issue/Meet-145/Obsuzhdenie-privyazki-STZ-k-nestandartizirovannym-poziciyam-SPGZ-i-dorabotki-EAIST-s-celyu-predostavleniya-zakazchikam) Обсуждение привязки СТЗ к нестандартизированным позициям СПГЗ и доработки ЕАИСТ с целью предоставления заказчикам возможности использования в определённых случаях позиций СПГЗ, которые стали неактуальными или были удалены с момента формирования лота и до момента публикации извещения (Расширение стандартизации и наследование КПГЗ)" meeting.
 
* This task is carrying out in the card "[DA-54](https://youtrack.gbuac.moscow/issue/DA-54/Podgotovit-analitiku-statistiku-izmenenij-KPGZ-SPGZ-v-2025-godu-v-razreze-osnovanij) Подготовить аналитику/статистику изменений КПГЗ/СПГЗ в 2025 году в разрезе оснований"

* The results of this task will be used in the following tasks:

  - [SGL-105](https://youtrack.gbuac.moscow/issue/SGL-105/Soglasovat-FT-na-Realizaciyu-ispolzovaniya-arhivnyh-KPGZ-i-SPGZ-zapros-ot-GRBS) Согласовать ФТ на Реализацию использования архивных КПГЗ и СПГЗ (запрос от ГРБС)

  - [Meet-145](https://youtrack.gbuac.moscow/issue/Meet-145/Obsuzhdenie-privyazki-STZ-k-nestandartizirovannym-poziciyam-SPGZ-i-dorabotki-EAIST-s-celyu-predostavleniya-zakazchikam) Обсуждение привязки СТЗ к нестандартизированным позициям СПГЗ и доработки ЕАИСТ с целью предоставления заказчикам возможности использования в определённых случаях позиций СПГЗ, которые стали неактуальными или были удалены с момента формирования лота и до момента публикации извещения (Расширение стандартизации и наследование КПГЗ)
  
### **1.1.2. Key stakeholders**

The key stakeholders of this task are following:

* Department of Economic Policy and Development of the City of Moscow

* Department of Competition Policy of the City of Moscow

* Moscow city customers

### **1.1.3. SMART questions**

The result of completing this task should be the answers to the following questions

|**N**|**Type**|**Question**   |
|-|-----|---------------|
| 1| - Specific <br> - Measurable <br> - Relevant <br> - Time-bound <br>  | How often  did the SPGZ change in 2025 for each of the scenarios (cases) described in "Матрица изменений" file, prepared during the work on task "[IS_Task-318](https://youtrack.gbuac.moscow/issue/IS_Task-318/Sformirovat-tablicu-izmenenij-SPGZ-posle-kotoryh-trebuetsya-aktualizaciya-lota) Сформировать таблицу изменений СПГЗ, после которых требуется актуализация лота".|
| 2| ... | ... |

### **1.1.4. Scope of work (SOW)**

|N | Deliverables | Timeline | Milestones | Results    | What SMART questions will be answered |
|-|-----------   | ----     |------------|------------|-----------------------------|
|1 | Dataset with information about changes SPGZ in 2025 from Catalog database | 18.12.2025 | Completion of extracting data from Catalog database | Dataset with information about changes SPGZ in 2025 from Catalog database | - |
|2 | Clean dataset | 19.12.2025 | Dataset is clean and usable | - |
|3 | Exploratory data analysis (EDA) | 21.12.2025 | Completion of EDA - understanding of dataset is formed | EDA highlights room for investigation. Questions were formulated based on the results of EDA. | - |
|4 | Engineered features in the form of new columns| 22.12.2025 | New columns engineered to provide more accessible information |
|5 | Statistical insights | 24.12.2025 | Completion of statistical analysis | Statistical analysis conducted | How often  did the SPGZ change in 2025 for each of the scenarios (cases) described in "Матрица изменений" file, prepared during the work on task "[IS_Task-318](https://youtrack.gbuac.moscow/issue/IS_Task-318/Sformirovat-tablicu-izmenenij-SPGZ-posle-kotoryh-trebuetsya-aktualizaciya-lota) Сформировать таблицу изменений СПГЗ, после которых требуется актуализация лота". |
|6 | Visual aids | 25.12.2025 | Completion of visualizations | Visualizations created | - |
|7 | Final report| 27.12.2025 | Final report submission | Final report completed | - |


## **1.2 Guiding questions**

| N | Question  | Answer     |
|-|-----------|------------|
| 1 | What problem need to be solved using the result of this task | The problem that needs to be solved is that customers need to make changes in they lots if the SPGZ positions used in those lot's specifications are changed or deleted |
| 2 | How can the insights (results) of this task drive business decisions | This analysis insights will be used to determine necessary EAIST upgrades and their priorities.|
<br>
<br>
<br>

# **2. Data preparation**

## **2.1. Key tasks and delivirables**

### **2.1.1. Collecting the necessary data**

Data was collected from several Catalog database tables using the SQL query "spgz_kpgz_relations_packages.sql".

### **2.1.2. Downloading and storing data**

The data was downloaded and stored at the enterprise cloud and available is at this [link](https://drive.ac-mos.ru/f/4038789)

```{r, warning=FALSE, message=FALSE}
# df <- read.csv("C:/Users/GudievZK/Nextcloud/Data analysis/DA-54/2025.12.18_spgz_kpgz_ktru_pack_rels_db_ktlg.csv", fileEncoding = "Windows-1251")
df <- read_excel("/Users/zelimkhan/Desktop/DA-54/2025.12.18_spgz_kpgz_ktru_pack_rels_db_ktlg.xlsx", guess_max = 60000)
df_src <- read_excel("/Users/zelimkhan/Desktop/DA-54/2025.12.18_spgz_kpgz_ktru_pack_rels_db_ktlg.xlsx", guess_max = 60000)
```


### **2.1.3. Data organization**

The data is organized as a table (in csv format) containing **`r nrow(df)`** rows. The rows contain information about versions of SPGZ positions which were changed or deleted in 2025.

### **2.1.4. Data credibility determination**

The ROCCC framework and Data integrity methodology were used to determine the data credibility

##### **ROCCC framework**

| N | ROCCC Dimension |  Description  | Conclusion | Comments  |
|-  |      -----------|-----------    | -----------|-----------|
|1  | Reliable | Source is dependable and consistent. | See to the conclusion below the table | It depends on ROCCC dimensions listed below |
|2  | Original | The data obtained from the original source (or if the data is not obtained from the original source, it is possible compere and verify the data with the original source) | The data obtained from the original source (Catalog database).
|3 | Comprehensive | Covers all necessary cases and variables | The data covers all necessary cases and variables | |
|4 | Current | The data source aren't out of date or irrelevant | The data is current and contains the necessary information for 2025 year | | 
|5 | Cited   | The data source is cited and vetted | The data source is cited and veted | |
**Conclusion on reliability of the data source based on the on ROCCC framework: The data source is reliable.**

##### **Data integrity**
| N | Data integrity dimension | Description | Conclusion |
|-  |       -----------|-----------    | -----------|
|1  |  Accuracy | The degree to which the data conforms to the actual entity being measured or described | The data conforms actual changes in the ZPGZ in 2025 |
|2  | Completeness | Data completeness is a dimension of data quality that measures whether all necessary or required data is present and available for dataset to fulfill to intended purpose.| The dataset contains all necessary cases and variables.|
|3  | Consistency | Data follows the same formats and rules.| The date are consistency |
|4  | Trustworthiness | The trustworthiness refers to the degree to which information can be relied upon as accurate, credible and secure, allowing users confidently make decisions and conduct analysis based on it.  | The data is trustworthiness |
**Conclusion on credibility of the data source based on Data integrity methodology: The source is credibility.**

#### **2.1.5. Bias**
The dataset come from Catalog database and reflect actaull changes in the SPGZ in 2025

#### **2.1. 6. Conclusion on the admissibility of using data**
The data is credible and can be use for analysis

## 2.2 Guiding questions
| N | Questions |  Answers  |
|-  |-----------|-----------|
|1  | Where the data is located? | The data was downloaded from Catalog database and stored in the enterprise cloud at this [link](https://drive.ac-mos.ru/f/4038789) (see section "2.1.2. Downloading and storing data")|
|2  | How is the data organized? | The data is organized as a table consisting of **`r nrow(df)`** and **`r ncol(df)`** |
|3  | Are there issues with bias or credibility in the data? Does the data ROCCC? | The data is unbiased and credible. Data is integrity and meets the ROCCC methodology requirements (see section the "2.1.4. Data credibility determination" section)|
|4  | How are issues of licencing, privacy, security, and accessibility addressed? | 


<br>
<br>
<br>

# **3. Data processing**

## Stage 1 of EDA: Data Discovery & Profiling

Table overview 
```{r}
df
```

Clean names
```{r}
# toString(names(df))
df <- clean_names(df)
```

Columns names
```{r}
names(df)
```
Number of rows and columns
```{r}
dim(df)
```
Data types
```{r}
glimpse(df)
```
Data classes
```{r}
sapply(df, class)
```
Checking duplicates. Number of duplicate columns in each column.
```{r}
sapply(df, function(x) sum(duplicated(x)))
```
Checking uniqueness. 

Number of unique rows.
```{r}
df %>% n_distinct()
```

Number of unique values in each column
```{r}
sapply(df, function(x) sum(n_distinct(x)))
```

Checking completness
```{r}
colSums(is.na(df))
```

**Converting data types**
Converting all columns with fewer than 10 unique (non-NA) values into factors

```{r}
# df <- df %>% 
#   mutate(
#     across(
#       where(~ n_distinct(.x, na.rm = TRUE) < 10),
#       as.factor
#     )
#     )
```

Converting variables with date and time values from character format to date format.
```{r}
df <- df %>% 
  mutate(
    across(
      contains("date_") & where(is.character), 
           ~ as.Date(ymd_hms(.x)))
  )

```

Filtering data
```{r}
df <- df %>% 
  filter(
    (isexcluded_spgz == 0 & !is.na(package_user_name)) & # deleting excluded spgz and spgz without author (which where created by mistake)
      (
           year(editdate_spgz) == 2025 |
         year(enddate_version_spgz) == 2025)
        )
```

Deleting duplicate rows
```{r}
df <- df %>% 
   distinct()
```

Deleting unnecessary columns
```{r}
df <- df %>% 
  select(id_entity_spgz_katalog, id_version_spgz_katalog, name_spgz, isstandardized_spgz, isproject_spgz, isdeleted_spgz, 
  ispcp_spgz, createdate_spgz, editdate_spgz, begindate_version_spgz, enddate_version_spgz, okpd2_code_spgz, 
  okpd2_name_spgz, ktru_code_spgz, ktru_name_spgz, rels_old_pgz_id, rels_linktype, rels_new_pgz_id, rels_is_actual, 
  rels_source_pgz_id, package_name, package_user_name, package_reason_name, id_entity_kpgz, name_kpgz, code_kpgz)

# toString(names(df))
```


```{r}
dups <- df %>%
  group_by(across(everything())) %>%  # group by all columns
  filter(n() > 1) %>%                 # keep groups with more than one row
  ungroup()
```

**Adding new columns**
Adding a column with number of values for each id_entity_spgz_katalog
```{r}
df <- df %>% 
  group_by(id_entity_spgz_katalog) %>% 
  mutate(
    n_id_entity_spgz_katalog = n()) %>% 
  ungroup()
```

Adding a column with the numbers of unique values id_version_spgz_katalog per id_entity_spgz_katalog
```{r}
df <- df %>% 
  group_by(id_entity_spgz_katalog) %>% 
  mutate(
    n_unique_id_version_cataloge = n_distinct(id_version_spgz_katalog, na.rm = TRUE)) %>% 
      ungroup()
```

Adding column with difference between "id_entity_spgz_katalog" and "id_version_spgz_katalog"
```{r}
df <- df %>% 
  mutate(
    diff_id_entity_id_version = n_id_entity_spgz_katalog - n_unique_id_version_cataloge
  )
```

Adding a column with deletion information
```{r}
df <- df %>% 
  group_by(id_entity_spgz_katalog) %>% 
  mutate(
    is_deleted_spgz = as.integer(sum(as.integer(isdeleted_spgz), na.rm = TRUE) > 1)
    ) %>% 
      ungroup()
```
Adding column with information about relation...


Adding column with information about standardizing in 2025
```{r}
df <- df %>% 
  group_by(id_entity_spgz_katalog) %>% 
  mutate(
    was_standartized_25 = as.integer(
      n_distinct(trimws(as.character(isstandardized_spgz)), na.rm = TRUE) > 1
      )
    ) %>% 
  ungroup()

```








dssd
```{r}
# rels_old_pgz_id <- df %>% filter(!is.na(rels_old_pgz_id)) %>% select(rels_old_pgz_id)
# df_rels <- df %>% filter(id_version_spgz_katalog %in% rels_old_pgz_id)

df %>% filter(
  id_version_spgz_katalog == "30425543" |
    rels_old_pgz_id == "30425543" |
    id_entity_spgz_katalog == "29270927" |
    id_entity_spgz_katalog == "30424361" |
    id_version_spgz_katalog == "30425542"
) %>% write_xlsx("/Users/zelimkhan/Desktop/DA-54/2025.12.21_df_0.xlsx")
```

```

<br>
<br>
<br>

# **4. Data analysis**
<br>
<br>
<br>

# **5. Report**
<br>
<br>
<br>

# **6. Usage result**

```{r}
write_xlsx(df, "/Users/zelimkhan/Desktop/DA-54/2025.12.21_df.xlsx")
```

:::