---
title: "DA 54. Analysis of changes to the SPGZ in 2025 in terms of grounds"
author: "Zelimkhan Gudiev"
date: "2025-12-17"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: TRUE
---

::: {style="text-align: justify"}

# Install packages and attach libraries
```{r setup, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
# install.packages("psych")
# install.packages("lubridate")
# install.packages("purrr")
# install.packages("GGally")
# install.packages("skimr")
# install.packages("Writexl")
# install.packages("corrplot")
library(psych)
library(googlesheets4)
library(tidyr)
library(dplyr)
library(janitor)
library(ggplot2)
library(lubridate)
library(purrr)
library(GGally)
library(readxl)
library(skimr)
library(writexl)
library(corrplot)
```

# **1. Business task, context and scope of work**

## **1.1. Key tasks and deliverable(s)**

### **1.1.1. Business task**

Determine the number of changes made to the SPGZ in 2025 for each of possible scenarios (cases) for the subsequent determination of frequency of scenarios (cases) in which customers receive notifications (in accordance with "[EAIST_JB-8765](https://pm.mos.ru/issue/EAIST_JB-8765) Реализовать уведомления пользователям при изменении позиций КПГЗ/СПГЗ, которые были запланированы в лотах") and scenarios (cases) that entail the need for customers to make changes to the lot specifications. 

**Notes**

* The assignment to carry out of this task was issued at the "[Meet-145](https://youtrack.gbuac.moscow/issue/Meet-145/Obsuzhdenie-privyazki-STZ-k-nestandartizirovannym-poziciyam-SPGZ-i-dorabotki-EAIST-s-celyu-predostavleniya-zakazchikam) Обсуждение привязки СТЗ к нестандартизированным позициям СПГЗ и доработки ЕАИСТ с целью предоставления заказчикам возможности использования в определённых случаях позиций СПГЗ, которые стали неактуальными или были удалены с момента формирования лота и до момента публикации извещения (Расширение стандартизации и наследование КПГЗ)" meeting.
 
* This task is carrying out in the card "[DA-54](https://youtrack.gbuac.moscow/issue/DA-54/Podgotovit-analitiku-statistiku-izmenenij-KPGZ-SPGZ-v-2025-godu-v-razreze-osnovanij) Подготовить аналитику/статистику изменений КПГЗ/СПГЗ в 2025 году в разрезе оснований"

* The results of this task will be used in the following tasks:

  - [SGL-105](https://youtrack.gbuac.moscow/issue/SGL-105/Soglasovat-FT-na-Realizaciyu-ispolzovaniya-arhivnyh-KPGZ-i-SPGZ-zapros-ot-GRBS) Согласовать ФТ на Реализацию использования архивных КПГЗ и СПГЗ (запрос от ГРБС)

  - [Meet-145](https://youtrack.gbuac.moscow/issue/Meet-145/Obsuzhdenie-privyazki-STZ-k-nestandartizirovannym-poziciyam-SPGZ-i-dorabotki-EAIST-s-celyu-predostavleniya-zakazchikam) Обсуждение привязки СТЗ к нестандартизированным позициям СПГЗ и доработки ЕАИСТ с целью предоставления заказчикам возможности использования в определённых случаях позиций СПГЗ, которые стали неактуальными или были удалены с момента формирования лота и до момента публикации извещения (Расширение стандартизации и наследование КПГЗ)
  
### **1.1.2. Key stakeholders**

The key stakeholders of this task are following:

* Department of Economic Policy and Development of the City of Moscow

* Department of Competition Policy of the City of Moscow

* Moscow city customers

### **1.1.3. SMART questions**

The result of completing this task should be the answers to the following questions

|**N**|**Type**|**Question**   |
|-|-----|---------------|
| 1| - Specific <br> - Measurable <br> - Relevant <br> - Time-bound <br>  | How often  did the SPGZ change in 2025 for each of the scenarios (cases) described in "Матрица изменений" file, prepared during the work on task "[IS_Task-318](https://youtrack.gbuac.moscow/issue/IS_Task-318/Sformirovat-tablicu-izmenenij-SPGZ-posle-kotoryh-trebuetsya-aktualizaciya-lota) Сформировать таблицу изменений СПГЗ, после которых требуется актуализация лота".|
| 2| ... | ... |

### **1.1.4. Scope of work (SOW)**

|N | Deliverables | Timeline | Milestones | Results    | What SMART questions will be answered |
|-|-----------   | ----     |------------|------------|-----------------------------|
|1 | Dataset with information about changes SPGZ in 2025 from Catalog database | 18.12.2025 | Completion of extracting data from Catalog database | Dataset with information about changes SPGZ in 2025 from Catalog database | - |
|2 | Clean dataset | 19.12.2025 | Dataset is clean and usable | - |
|3 | Exploratory data analysis (EDA) | 21.12.2025 | Completion of EDA - understanding of dataset is formed | EDA highlights room for investigation. Questions were formulated based on the results of EDA. | - |
|4 | Engineered features in the form of new columns| 22.12.2025 | New columns engineered to provide more accessible information |
|5 | Statistical insights | 24.12.2025 | Completion of statistical analysis | Statistical analysis conducted | How often  did the SPGZ change in 2025 for each of the scenarios (cases) described in "Матрица изменений" file, prepared during the work on task "[IS_Task-318](https://youtrack.gbuac.moscow/issue/IS_Task-318/Sformirovat-tablicu-izmenenij-SPGZ-posle-kotoryh-trebuetsya-aktualizaciya-lota) Сформировать таблицу изменений СПГЗ, после которых требуется актуализация лота". |
|6 | Visual aids | 25.12.2025 | Completion of visualizations | Visualizations created | - |
|7 | Final report| 27.12.2025 | Final report submission | Final report completed | - |


## **1.2 Guiding questions**

| N | Question  | Answer     |
|-|-----------|------------|
| 1 | What problem need to be solved using the result of this task | The problem that needs to be solved is that customers need to make changes in they lots if the SPGZ positions used in those lot's specifications are changed or deleted |
| 2 | How can the insights (results) of this task drive business decisions | This analysis insights will be used to determine necessary EAIST upgrades and their priorities.|


# **2. Data preparation**

## **2.1. Key tasks and delivirables**

### **2.1.1. Collecting the necessary data**

Data was collected from several Catalog database tables using the SQL query "spgz_kpgz_relations_packages.sql".

### **2.1.2. Downloading and storing data

The data was downloaded and stored at the enterprise cloud and available is at this [link](https://drive.ac-mos.ru/f/4038789)

```{r}
df <- read.csv("C:/Users/GudievZK/Nextcloud/Data analysis/DA-54/2025.12.18_spgz_kpgz_ktru_pack_rels_db_ktlg.csv", fileEncoding = "Windows-1251")
```


### **2.1.3. Data organization**

The data is organized as a table (in csv format) containing **`r nrow(df)`** rows. The rows contain information about versions of SPGZ positions which were changed or deleted in 2025.

### **2.1.4 Data credibility determenation**

The ROCCC framework and Data integrity methodology were used to determine the data credibility

#### **2.1.4.1 ROCCC framework**

| N | ROCCC Dimension |  Description  | Conclusion | Comments  |
|-  |      -----------|-----------    | -----------|-----------|
|1  | Reliable | Source is dependable and consistent. | See to the conclusion below the table | It depends on ROCCC dimensions listed below |
|2  | Original | The data obtained from the original source (or if the data is not obtained from the original source, it is possible compere and verify the data with the original source) | The data obtained from the original source (Catalog database).
|3 | 

# 3. Data processing



# 4. Data analysis

# 5. Report

# 6. Usage result

:::