---
title: "Lettuce_growth_days"
author: "Zelimkhan"
date: "2025-09-22"
output: 
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---

# 0. Install packages and attach libraryes
```{r setup, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
# install.packages("psych")
# install.packages("lubridate")
# install.packages("purrr")
# install.packages("GGally")
library(psych)
library(googlesheets4)
library(tidyr)
library(dplyr)
library(janitor)
library(ggplot2)
library(lubridate)
library(purrr)
library(GGally)
library(readxl)
```
# Step 1. Ask
## 1.1 Guiding questions
::: {style="text-align: justify"}
| N | Questions |  Answers  |
|-  |-----------|-----------|
| 1 | What is the problem you are trying to solve?   | The challenge lies in understanding (finding patterns) how environmental conditions influence lettuce growth time. This knowledge  will help inform decisions to optimize growth cycle. | 
| 2 | How can you insights drive business decisions? | Knowing about environmental conditions that impact lettuce growth can help managers decide on construction of an appropriate greenhouse |

## 1.2 Key tasks
| N | Tasks     |  Results  |
|-  |-----------|-----------|
| 1 | Identify business task     | Analyze data (lettuce dataset) to understand how environmental conditions affect lettuce growth time. |
| 2 | Consider key stakeholders  | Managers |
| 3 | Create scope of work (SOW) | SOW is created |

## 1.3 Deliverable(s)

### 1.3.1 A clear statement of the business task
Analyze environmental conditions' impact on lettuce growth (finding patterns) and provide actionable insights for optimizing growth cycles.

### 1.3.2 Scope of work (SOW)
| N | Deliverables | Timeline | Milestones | Results |
|-  |-----------   | ----     |------------| -----------|
| 1 | Clean dataset | Day 1 | Completion of data download and preparation | Dataset is clean and usable |
| 2 | Exploratory data analysis (EDA) findings (form: spreadsheet, google doc summary) | Day 2 | Completion of EDA - understanding of dataset is formed | EDA highlights room for investigation. Questions were formulated based on the results of EDA. |
| 3 | Engineered features in the form of new columns (examples may include Avg 3 day growth, Growth since day before, etc.) | Day 3 | Completion of feature engineering | New columns engineered to provide more accessible information | ... |
| 4 | Statistical insights | Day 3 | Completion of statistical analysis | Statistical analysis conducted |
| 5 | Visual Aids (images, dashboards) | Days 4-5 | Completion of visualizations | Visualizations created | 
| 6 | Final report (spreadsheet and PowerPoint) | Day 6-10 | Final report submission | Final report completed |
:::


# 2. Prepare
## 2.1 Guiding questions
| N | Questions |  Answers  |
|-  |-----------|-----------|
| 1 | Where is your data located?                                                   | The data is publicly available in Kaggle website at this [link](https://www.kaggle.com/datasets/jurijsruko/lettuce) |
| 2 | How is the data organized?                                                    | The data is organized as a table containing data on 70 plants. The rows contain environmental conditions for each of the 70 plants on each day throughout the growing season. The columns contain information about the plant identifier (Plant_ID) and the following environmental conditions, which were recorded daily for each of the 70 plants throughout the growing season: Date; Temperature (°C); Humidity (%); TDS Value (ppm); pH Level; Growth Days.|
| 3 | Are there issues with bias or credibility in this data? Does your data ROCCC? | 1) The lettuce growth dataset shows no major issues with bias or credibility, but as with any sample dataset, it may have some limitations: <br> 1.1) **Bias**: The dataset appears to come from a controlled educational or reaserch environment, so it may not present all real world farming conditions or varieties of lettuce. However, since it is used for analytical practice rather than policy or production decisions, the risk of bias affecting conclusions is minimal. <br> 1.2) **Credibility**: The data structure, values, and relationships appear internal consistent and realistic. Variable names, measurements units, and record counts align with exepted patterns, which supports the datasets credibility. <br> 2) **ROCCC** <br> 2.1) **Relaible.** The data is reliable (accurate, complete and unbiased). <br> 2.2) **Original.** The data is original. We can locate the original souce of data and we don't rely solely on third-party information. The dataset is original to the learning environment and was prepared for educational analysis by verified sources (Kaggle). <br> 2.3) **Comprehensive.** It includes all key variables needed to explore lettuce growth duration and related conditions, though it is simplified for training purposes. <br> 2.4) **Current.** The data sources aren't out of date or irrelivant. This data source is respected an their refresh their data reguraly. The dataset reflects a modern data structure and is suitable for current analytical tools (R, Excel, etc.), even if the measurements are not tied to a specific recent year. <br> 2.5) **Cited.** The data was obtaines from the Kaggle website. The dataset and its source are cited. <br>|
| 4 | How are you addressing licensing, privacy, security, and accessibility?       |-----------|
| 5 | How did you verify the data’s integrity?                                      |-----------|
| 6 |	How does it help you answer you questions?                                    |-----------|
| 7 | Are there any problems with the data?                                         |-----------|

## 2.2 Key tasks
## 2.3 Deliverables

## 2.1 Read data
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# df_0 <- read_sheet("https://docs.google.com/spreadsheets/d/1esw9BCXZDO1h4hwuaKowia5eVmflEKTS6NiouamSYA0/edit?gid=723124268#gid=723124268", sheet = "source_lettuce_dataset")
df_0 <- read_excel("C:/Users/GudievZK/Desktop/GitHub/Google Data Analytics Professional Certificate/1. Ask_Questions_to_Make_Data_Driven_Decisions/lettuce_dataset_SOW.xlsx", sheet = "source_lettuce_dataset")
```

```{r}
df_0
```

# 3. Exploratory data analysis (EDA)
## 3.1 Understand the dataset
What variables (columns) exist? 
```{r}
colnames(df_0)
```

What types of data do they contain (numeric, categorical, dates)?
```{r}
glimpse(df_0)
```
How many rows are there?
```{r}
dim(df_0)
```

## 3.2 Check data quality
Are there missing values?
```{r}
colSums(is.na(df_0))
```
Are there any duplicates or errors?
```{r}
sapply(df_0, function(x) sum(duplicated(x)))
```
Are values in a reasonable range (e.g., temperature not 500Â°C)?
```{r}
df_0 %>% select(-Date) %>% summary()
```
```{r}
df_0 <- df_0 %>% 
  mutate(
    # 1) Flatten: make every element a clean character string
    Date = map_chr(Date, ~ if (inherits(.x, "POSIXt"))
      format(.x, "%Y-%m-%d") else as.character(.x)),
    Date = trimws(Date),
    Date = gsub("[\u2010\u2011\u2012\u2013\u2014\u2212]", "-", Date), # normalize dashes
    Date = na_if(Date, ""),
    
    # 2) Parse: try ISO (Y-m-d) then US (m/d/Y); keep first non-NA
    Date = coalesce(
      ydm(Date, tz = "UTC", quiet = TRUE),
      mdy(Date, tz = "UTC", quiet = TRUE)
    )
  ) %>% arrange(Date)
```
Clean names 
```{r}
df <- clean_names(df_0)
```

## 3.3 Summarize statistics
Data frame summary
```{r}
summary(df)
```
All statistics by plant id
```{r}
all_stats_by_id <- df %>% group_by(plant_id) %>% 
  summarise(
            date_min = min(date),
            date_mean = mean(date),
            date_median = median(date),
            date_max = max(date),
            temperature_c_min = min(temperature_c),
            temperature_c_mean = mean(temperature_c),
            temperature_c_median = median(temperature_c),
            temperature_c_sd = sd(temperature_c),
            temperature_c_max = max(temperature_c),
            humidity_percent_min = min(humidity_percent),
            humidity_percent_mean = mean(humidity_percent),
            humidity_percent_median = median(humidity_percent),
            humidity_percent_sd = sd(humidity_percent),
            humidity_percent_max = max(humidity_percent),
            tds_value_ppm_min = min(tds_value_ppm),
            tds_value_ppm_mean = mean(tds_value_ppm),
            tds_value_ppm_median = median(tds_value_ppm),
            tds_value_ppm_sd = sd(tds_value_ppm),
            tds_value_ppm_max = max(tds_value_ppm),
            p_h_level_min = min(p_h_level),
            p_h_level_mean = mean(p_h_level),
            p_h_level_median = median(p_h_level),
            p_h_level_sd = sd(p_h_level),
            p_h_level_max = max(p_h_level),
            growth_days_min = min(growth_days),
            growth_days_mean = mean(growth_days),
            growth_days_median = median(growth_days),
            growth_days_sd = sd(growth_days),
            growth_days_max = max(growth_days)
  ) %>% 
  mutate(
    across(where(is.numeric), ~ round(.x, 2))
         )
all_stats_by_id
```
Summary of all statistics by plant id
```{r}
all_stats_by_id %>% summary()
```
Statistics (basic) by plant id
```{r}
all_stats_by_id %>% 
  select(plant_id, temperature_c_mean,
         humidity_percent_mean,
         tds_value_ppm_mean,
         p_h_level_mean,
         growth_days_max)
```

Statistics by maximum growth day
```{r}
df %>% group_by(plant_id) %>% 
  mutate(growth_days_max = max(growth_days)) %>% 
  group_by(growth_days_max) %>% 
  summarise(
    plant_id_n = n_distinct(plant_id),
            temperature_c_mean = mean(temperature_c),
            humidity_percent_mean = mean(humidity_percent),
            tds_value_ppm_mean = mean(tds_value_ppm),
            p_h_level_mean = mean(p_h_level)
  ) %>% 
  mutate(
    across(
      where(is.numeric), ~ round(.x, 2)
      )
  )
```

# 3.4 Visualize relationships

Histograms (distribution)
```{r}
df %>% select(where(is.numeric)) %>% 
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "value") %>% 
  ggplot(aes(x = value, fill = variable)) +
  geom_histogram(bins = 30, alpha = 0.6, color = "black", na.rm = TRUE) +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
# theme_minimal() + 
  guides(fill = "none")
  
```

Boxplots (outliers).
```{r}
df %>% select(where(is.numeric)) %>% 
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "value") %>% 
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(outliers.alpha = 0.35) +
  labs(x = NULL, y = NULL, title = "Boxplots of numeric variables") +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
  # theme_minimal() +
  guides(fill = "none")
  
```

Time-series plots (trends over time)

```{r}
df %>%
  pivot_longer(-c(date, plant_id), names_to = "variable", values_to = "value") %>% 
  group_by(variable, date) %>%
  summarise(
    mean = mean(value, na.rm = TRUE),
    p25  = quantile(value, 0.25, na.rm = TRUE),
    p75  = quantile(value, 0.75, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(variable, date) %>% 
  ggplot(aes(date, mean)) +
  geom_ribbon(aes(ymin = p25, ymax = p75), alpha = 0.15) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_y", ncol = 2) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  labs(x = NULL, y = NULL, title = "Daily mean with IQR band") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.minor = element_blank())
```
c
```{r}
df %>% 
  pivot_longer(-c(date, plant_id), names_to = "variable", values_to = "values") %>% 
  group_by(variable, date) %>% 
  summarise(mean = mean(values)) %>% 
  arrange(variable, date) %>% 
  ggplot(aes(x = date, y = mean)) +
  geom_line() + 
  facet_wrap(~ variable, scales = "free_y", ncol = 2) + 
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  labs(x = NULL, y = NULL, title = "Daily mean") + 
  # theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.minor = element_blank())
```
Scatterplots (check correlation)
```{r}
df %>% select(where(is.numeric)) %>% 
  select(- plant_id) %>% 
  ggpairs(progress = FALSE,
          upper = list(continuous = wrap("cor", method = "pearson")),
          diag = list(continuous = "densityDiag"),
          lwer = list(continuous = wrap("points", alpha = 0.4, size = 0.7))
    
  )
```

# 3.5 Form hypotheses

Does higher humidity improve growth?

Are weekly growth patterns consistent?

Is there seasonality in the data?




