---
title: "EDA"
author: "Zelimkhan Gudiev"
date: "2026-03-12"
output: html_document
---

# 1. SETUP & LOAD DATA

Install required packages, load the diamonds dataset, and understand what we are working with before touching any statistics.

📦 Dataset — diamonds (ggplot2)
The diamonds dataset contains measurements of 53,940 diamonds. It includes price, carat weight, cut quality, color grade, clarity grade, and physical dimensions (x, y, z depth). It is a rich dataset with both numeric and categorical variables — perfect for a real-world workflow.

## INSTALL PACKAGES (run once, then comment out) ##
All packages used throughout this entire workflow
```{r}
# install.packages(c(
#   "tidyverse",    # includes dplyr, ggplot2, tidyr, readr, purrr
#   "psych",        # describe() — rich descriptive statistics table
#   "moments",      # skewness() and kurtosis() functions
#   "corrplot",     # correlation matrix heatmap visualization
#   "naniar",       # visualize missing values patterns
#   "patchwork",    # combine multiple ggplot charts into one figure
#   "scales"        # formatting axis labels (percent, comma, dollar)
# ))
```


## LOAD LIBRARIES ##
```{r}
library(tidyverse)   # loads dplyr, ggplot2, tidyr, etc. all at once
library(psych)       # describe()
library(moments)     # skewness(), kurtosis()
library(corrplot)    # corrplot()
library(naniar)      # vis_miss(), miss_var_summary()
library(patchwork)   # plot1 + plot2 + plot3 layout
library(scales)      # dollar(), percent(), comma()
```

## LOAD DATASET ##
# diamonds is built into ggplot2, so no file to import

```{r}
df <- diamonds    # assign to 'df' so our code is generalisable
```

## ── GLOBAL PLOT THEME ─────────────────────────────────────────
Set once here so ALL charts look consistent throughout
```{r}
theme_set(theme_minimal(base_size = 12) +
            theme(
              plot.title    = element_text(face = "bold", size = 14),
              plot.subtitle = element_text(color = "gray50", size = 11),
              panel.grid.minor = element_blank()  # remove minor gridlines
            )
)
```

